---
title: "UNHCR"

#output: html_document
output: 
  flexdashboard::flex_dashboard:
    orientation: row
    vertical_layout: fill
runtime: shiny

---

```{r setup, include=FALSE}
library(flexdashboard)#https://pkgs.rstudio.com/flexdashboard/
library(readr)
library(tidyverse)
#library(dplyr)
#library(plotly)
library(crosstalk)
library(dygraphs)
library(lubridate)
#library(xts)
library(shiny)
library(knitr)
#library(DT)
library(highcharter)

#https://jkunst.com/highcharter/
#https://api.highcharts.com/highcharts/
#https://jkunst.com/blog/posts/2020-06-26-valuebox-and-sparklines/

hc_theme_sparkline_vb <- function(...) {
  
  theme <- list(
    chart = list(
      backgroundColor = NULL,
      margins = c(0, 0, 0, 0),
      spacingTop = 0,
      spacingRight = 0,
      spacingBottom = 0,
      spacingLeft = 0,
      plotBorderWidth = 0,
      borderWidth = 0,
      style = list(overflow = "visible")
    ),
    xAxis = list(
      visible = FALSE, 
      endOnTick = FALSE, 
      startOnTick = FALSE
    ),
    yAxis = list(
      visible = FALSE,
      endOnTick = FALSE, 
      startOnTick = FALSE
    ),
    tooltip = list(
      outside = FALSE,
      shadow = FALSE,
      borderColor = "transparent",
      botderWidth = 0,
      backgroundColor = "transparent",
      style = list(textOutline = "5px white")
    ),
    plotOptions = list(
      series = list(
        marker = list(enabled = FALSE),
        lineWidth = 2,
        shadow = FALSE,
        fillOpacity = 0.25,
        color = "#FFFFFFBF",
        fillColor = list(
          linearGradient = list(x1 = 0, y1 = 1, x2 = 0, y2 = 0),
          stops = list(
            list(0.00, "#FFFFFF00"),
            list(0.50, "#FFFFFF7F"),
            list(1.00, "#FFFFFFFF")
          )
        )
      )
    ),
    credits = list(
      enabled = FALSE,
      text = ""
      )
    )
  
  theme <- structure(theme, class = "hc_theme")
  
  if (length(list(...)) > 0) {
    theme <- hc_theme_merge(
      theme,
      hc_theme(...)
    )
  }
  
  theme
}


valueBoxSpark <- function(value, title=NULL, sparkobj = NULL, subtitle=NULL, info = NULL, 
                          icon = NULL, color = "aqua", width = 4, href = NULL){
  
  shinydashboard:::validateColor(color)
  
  if (!is.null(icon))
    shinydashboard:::tagAssert(icon, type = "i")
  
  info_icon <- tags$small(
    tags$i(
      class = "fa fa-info-circle fa-lg",
      title = info,
      `data-toggle` = "tooltip",
      style = "color: rgba(255, 255, 255, 0.75);"
      ),

    class = "pull-right float-right"
    )

  boxContent <- div(
    class = paste0("small-box bg-", color),
    div(
      class = "inner",
      if (!is.null(sparkobj)) info_icon,
      h3(value),
      if (!is.null(sparkobj)) sparkobj,
      p(subtitle)
      ),
    if (!is.null(icon)) div(class = "icon-large icon", icon, style = "z-index; 0")
    )

  div(
    class = if (!is.null(width)) paste0("col-sm-", width), 
    boxContent
    )
}

```

```{r load_clean_data, echo=FALSE}

# download data using unhcrdatapackage 
#https://github.com/Edouard-Legoupil/unhcrdatapackage

population=unhcrdatapackage::end_year_population_totals
code=unhcrdatapackage::reference%>%
  select(iso_3,UNHCRBureau)
population=merge(population,code,by.x = "CountryAsylumCode",by.y = "iso_3",all.x = T)

# Let's shorten long country names in all columns
population[population == "United Kingdom of Great Britain and Northern Ireland"] <- "UK"
population[population == "Iran (Islamic Republic of)"] <- "Iran"
population[population == "United Republic of Tanzania"] <- "Tanzania"
population[population == "Venezuela (Bolivarian Republic of)"] <- "Venezuela"
population[population == "Republic of North Macedonia"] <- "North Macedonia"
population[population == "China, Hong Kong Special Administrative Region"] <- "China"
population[population == "Republic of Moldova"] <- "Moldova"
population[population =="Democratic Republic of the Congo"]<-"Congo DR"
population[population =="Syrian Arab Republic"]<-"Syria"
population[population =="United States of America"]<-"USA"
population[population =="Central African Republic"]<-"CAR"

population$CountryOriginName<-as.factor(population$CountryOriginName)
population$CountryOriginCode<-as.factor(population$CountryOriginCode)
population$CountryAsylumName<-as.factor(population$CountryAsylumName)
population$CountryAsylumCode<-as.factor(population$CountryAsylumCode)

sd_Trend<-SharedData$new(population)


```

# Overview 

Column {.sidebar}
--------------------------------------------------

Refugees, internally displaced persons, asylum-seekers and other persons of concern to UNHCR 

```{r}

filter_checkbox("UNHCRBureaue", "Select Bureau", sd_Trend, group = ~UNHCRBureau,
                               inline = FALSE)

sliderInput(inputId ="year", label = "Year:",
            min = 2000, max = 2021, value =2021, step = 1, sep = "") # use sep = "" to remove the ',' from the integer year

#checkboxGroupInput("POCs","",choices = c("Refugee","Asylum seeker","IDP","Stateless","Other of concern"),selected = "REF")


```

Data sources:
[www.unhcr.org/refugee-statistics/](https://www.unhcr.org/refugee-statistics/)

Draft!\\

Tadesse 

Row
-------------------------------------

### **Total persons of concern** {.value-box}

```{r}

data1 <- reactive({
  sd_Trend$data(withSelection = TRUE, withFilter = TRUE)%>% 
    filter(Year>1999)%>%
    select(Year,REF, ASY,IDP,STA,OOC)%>%
    group_by(Year)%>%
    summarise(R=sum(REF),
              Asylums=sum(ASY),
              IDPs=sum(IDP), 
              Stateless=sum(STA),
              OOCs=sum(OOC),T=sum(REF+ASY+IDP+OOC+STA))
})

 
renderValueBox({

  hc <- hchart(data1(), "area", hcaes(x=Year, y=T), name = "Summary",color="red")  %>%
     hc_size( width = 150, height = 50) %>%
  #hc_size(height = 100) %>%
  hc_credits(enabled = FALSE) %>%
  hc_add_theme(hc_theme_sparkline_vb())
  v0=data1()%>%filter(Year==input$year)
  v=paste(format(sum(v0$T), big.mark = ","), "", sep = " ")
  v=as.character(v)
  valueBoxSpark(
   value=v,
   sparkobj = hc,
   #icon = icon("code"),
   width = 20,
   color = "light-blue",
   href = NULL

    )
})


```

### Refugees

```{r}

renderValueBox({
  
  hc <- hchart(data1(), "area", hcaes(x=Year, y=R), name = "Refugees",color="red")  %>% 
     hc_size( width = 150, height = 50) %>%
  #hc_size(height = 100) %>% 
  hc_credits(enabled = FALSE) %>% 
  hc_add_theme(hc_theme_sparkline_vb()) 
  v0=data1()%>%filter(Year==input$year)
  v=paste(format(sum(v0$R), big.mark = ","), "", sep = " ")
  valueBoxSpark(
   value=v,
   sparkobj = hc,
   #subtitle =  "Refugees",
   #icon = icon("code"),
   width = 20,
   color = "light-blue",
   href = NULL
   
    )
})

```

### Asylum seekers

```{r}

renderValueBox({
  
  hc <- hchart(data1(), "area", hcaes(x=Year, y=Asylums), name = "Asylum seekers",color="red")  %>% 
     hc_size( width = 150, height = 50) %>%
    hc_credits(enabled = FALSE) %>% 
  hc_add_theme(hc_theme_sparkline_vb())
  
  v0=data1()%>%filter(Year==input$year)
  v=paste(format(sum(v0$Asylums), big.mark = ","), "", sep = " ")
  valueBoxSpark(
   value=v,
   sparkobj = hc,
   width = 20,
   color = "light-blue",
   href = NULL
   
    )
})


```

### Internally displaced persons

```{r}

renderValueBox({
  
  hc <- hchart(data1(), "column", hcaes(x=Year, y=IDPs), name = "IDPs",color="white")  %>% 
  hc_size( width = 150, height = 50) %>% 
  hc_credits(enabled = FALSE) %>% 
  hc_add_theme(hc_theme_sparkline_vb()) 
  v0=data1()%>%filter(Year==input$year)
  v=paste(format(sum(v0$IDPs), big.mark = ","), "", sep = " ")
  valueBoxSpark(
   value=v,
   sparkobj = hc,
   #icon = icon("code"),
   width = 30,
   color = "light-blue",
   href = NULL
   
    )
})

```

### Stateless
```{r}

renderValueBox({
  
  hc <- hchart(data1(), "column", hcaes(x=Year, y=Stateless), name = "Stateless",color="white")  %>% 
  hc_size( width = 150, height = 50) %>% 
  hc_credits(enabled = FALSE) %>% 
  hc_add_theme(hc_theme_sparkline_vb()) 
  v0=data1()%>%filter(Year==input$year)
  v=paste(format(sum(v0$Stateless), big.mark = ","), "", sep = " ")
  valueBoxSpark(
   value=v,
   sparkobj = hc,
   #icon = icon("code"),
   width = 30,
   color = "light-blue",
   href = NULL
   
    )
})

```

### Other of concerns
```{r}

renderValueBox({
  
  hc <- hchart(data1(), "column", hcaes(x=Year, y=OOCs), name = "Others",color="white")  %>% 
  hc_size( width = 150, height = 50) %>% 
  hc_credits(enabled = FALSE) %>% 
  hc_add_theme(hc_theme_sparkline_vb()) 
  v0=data1()%>%filter(Year==input$year)
  v=paste(format(sum(v0$OOCs), big.mark = ","), "", sep = " ")
  valueBoxSpark(
   value=v,
   sparkobj = hc,
   #icon = icon("code"),
   width = 30,
   color = "light-blue",
   href = NULL
   
    )
})

```

Row
-----------------------------------------------------------------------

### Displacement trend

```{r ,echo=FALSE}

data <- reactive({
  sd_Trend$data(withSelection = TRUE, withFilter = TRUE)%>% 
    select(Year,REF, ASY,IDP,STA,OOC)%>%
    group_by(Year)%>%
    summarise(R=sum(REF),
              Asylums=sum(ASY),
              IDPs=sum(IDP), 
              Stateless=sum(STA),
              OOC=sum(OOC))
})

renderDygraph(
  dygraph(data()%>%filter(
    Year>1989), main = "Persons of concern to UNHCR")%>%
  dySeries("R",  color = "red", label = "Refugees") %>% 
  dyOptions(axisLabelFontSize=10)%>%
    dyAxis(
    "y",axisLabelFontSize=10,
    #label = "Asylum seekers",
    #http://stackoverflow.com/questions/2901102/how-to-print-a-number-with-commas-as-thousands-separators-in-javascript
    valueFormatter = 'function(d){return d.toString().replace(/\\B(?=(\\d{3})+(?!\\d))/g, ",");}',
    axisLabelFormatter = 'function(d){return d.toString().replace(/\\B(?=(\\d{3})+(?!\\d))/g, ",");}'
  )%>%
  #dySeries( "Year",dyOptions(labelsKMB = FALSE))%>% 
  dyRangeSelector() %>%
  dyEvent("1994", "Rwanda Genocide", labelLoc = "bottom")%>%
  dyEvent("2003", "U.S.A invasion of Iraq", labelLoc = "bottom")%>%
  dyEvent("2011", "Syria crisis", labelLoc = "bottom")%>%
  dyEvent("2013", "South sudan civil war", labelLoc = "bottom")%>%
  dyEvent("2010", "Arab spring", labelLoc = "bottom")%>%
  dyEvent("2016", "Brexit", labelLoc = "bottom")%>%
  dyAnnotation("2015", text = "G", tooltip = "Germany allowed Syrian refugees into the country")%>%
  dyEvent("2020", "Nagorno-Karabakh clashes", labelLoc = "bottom")%>% 
  dyEvent("2021", "U.S.A withdrwal from Afghanistan", labelLoc = "bottom")%>%  
  dyGroup(c("Asylums", "IDPs","Stateless","OOC"), drawPoints = TRUE, color = c("blue", "green","black","Aqua"))%>%
  dyLegend(width = 600)%>%
  dyLegend(show = "follow") # show legend only when hover on the series 
  
  
 )


```

Row 
-------------------------------------

### Top displacement situations by country of origin

```{r}

data_COO <- reactive({

   sd_Trend$data(withSelection = TRUE, withFilter = TRUE)%>%
    filter(Year==input$year)%>%
    group_by(Year,COA=CountryAsylumName,COO=CountryOriginName)%>%
    summarise(Refs=sum(REF),
              Asylums=sum(ASY),
              IDPs=sum(IDP),
              Stateless=sum(STA),
              OOC=sum(OOC),
              T=sum(REF+ASY))

})

renderHighchart({
  
d<- data_COO()%>%select(COO,Refs,Asylums,T)%>%
  group_by(COO)%>%
  summarise(Refs=sum(Refs),Asylums=sum(Asylums),T=sum(T))%>%
  arrange(desc(T))%>%
  head(10)%>%
  mutate(COO =reorder(COO,T))

  hchart(d, 'bar', hcaes(x = COO, y = T),name = "Total")%>%
    #hc_colors("#0072BC")%>% 
    hc_xAxis(title = list(text = "")) %>%
    hc_yAxis(title = list(text = ""))%>%
    hc_plotOptions(
      bar = list(
      dataLabels = list(
        enabled = FALSE
        #,format = "{point.T}"

      )
    )
  )


  })

```

### Top Refugee situation by country of asylum

```{r}

data_COA2 <- reactive({

   sd_Trend$data(withSelection = TRUE, withFilter = TRUE)%>%
    filter(Year==input$year)%>%
    group_by(Year,COA=CountryAsylumName,COO=CountryOriginName)%>%
    summarise(Refs=sum(REF),
              Asylums=sum(ASY),
              IDPs=sum(IDP),
              Stateless=sum(STA),
              OOC=sum(OOC),
              T=sum(REF+ASY))
})

renderHighchart({
  ddd=data_COA2()%>%
  group_by(COA)%>%
  summarise(Refs=sum(Refs),Asylums=sum(Asylums),T=sum(T))%>%
  arrange(desc(T))%>%
  head(10)%>%
  mutate(COA =reorder(COA,T))

  highchart()%>%
  hc_chart(type = "bar") %>%
  hc_plotOptions(
    column = list(
    dataLabels = list(enabled = FALSE),enableMouseTracking = FALSE),
    series=list(stacking = "percent"))%>%
  hc_xAxis(categories =ddd$COA)%>%
  hc_series(list(name="Refugees",data=ddd$Refs),
            list(name="Asylum-seekers",data=ddd$Asylums)
            # ,list(name="Asylum-seekers",data=ddd$IDPs),
            # list(name="Asylum-seekers",data=ddd$Stateless),
            # list(name="Asylum-seekers",data=ddd$OOC)
            )%>%
  hc_yAxis(labels = list(format = "{value}%"), max = 100)
  
})

```

### Top IDP situation

```{r}

renderHighchart({
  
d<- data_COO()%>%select(COO,IDPs,T)%>%
  group_by(COO)%>%
  summarise(IDPs=sum(IDPs))%>%
  arrange(desc(IDPs))%>%filter(IDPs>0)%>%
  head(10)%>%
  mutate(COO =reorder(COO,IDPs))

  hchart(d, 'bar', hcaes(x = COO, y = IDPs),name = "IDPs")%>%
    hc_xAxis(title = list(text = "")) %>%
    hc_yAxis(title = list(text = ""))%>%
    hc_plotOptions(
      bar = list(
      dataLabels = list(
        enabled = FALSE
        #,format = "{point.T}"

      )
    )
  )


  })

```

# Solutions

Row 
-------------------------------------

Coming soon: UNHCR durable solutions data analysis.\
https://www.unhcr.org/solutions.html


```{r}

```


# Machine Learning

Row 
-------------------------------------

Can we write a model to predict  solutions ,displacements, statelessness ...etc 

# Notes:

Data used : unhcrdatapackage(from R Package)


Stat as of June 2021

Row 
-------------------------------------



